\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-26 appendices.tex
Fixed proof of (previously) Lemma 15 and made changes in structure regarding this lemma. Now one lemma (Lemma 12) is used in three different results.

\subsection{Proof of Lemma~\ref{lem:bound_intervals_delta} \label{proof:lem:bound_intervals_delta}}

%\begin{lem} \label{lem:bound_intervals}
%Let $f(x)$ be a log-concave, symmetric, and differentiable density such that $\eta(x)$ is strongly unimodal. For any $x_1 > \ldots > x_n \in \mathbb R$, 
%\begin{equation}
%\frac{ \left(  \sum_{k=1}^n (-1)^{k+1} f(x_k) \right)^2} 
%{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
%\leq \max_i \eta(x_i). \label{eq:lem_bound_intervals}
%\end{equation}
%In particular, if $|x_i|>\epsilon > 0$ for all $i=1,\ldots,n$, then 
%\[
%\frac{ \left(  \sum_{k=1}^n (-1)^{k+1} f(x_k) \right)^2} 
%{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
%\leq \eta(\epsilon) < 0. 
%\]
%\end{lem}

@@

\Delta_n \triangleq \Delta_n(x_1,\ldots,x_n) \triangleq  \sum_{k=1}^n (-1)^{k+1} F(x_k),

@@

so 

@@

\eta(x) =  \frac{  \left(\delta_1(x) \right)^2}{\Delta_1(x)
(1-\Delta_1(x))}
= \frac{  \left( f (x) \right)^2}{ F(x)\left(1-F(x)\right)^2}. 

@@

We use induction on $n \in \mathbb N$ to show that the LHS of \eqref{eq:lem_bound_intervals_delta} is bounded from above by $\max_i \eta(x_i)$. The case $n=1$ is trivial. Assume that 

@@

\frac{ \left( \delta_n \right)^2} 
{\Delta_n \left(1- \Delta_n \right) } \leq \max_i \eta(x_i)

@@

The maximal value of the LHS of \eqref{eq:lem_bound_intervals_proof} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 

@@

& g(x_1,\ldots, x_N) \triangleq  2 \log \delta_N -  \log \Delta_N -  \log \left(1 - \Delta_N  \right),

@@

 The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by

@@

\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} f'(x_k)}{\delta_N} -\frac{(-1)^{k+1} f(x_k)}{\Delta_N } + \frac{(-1)^{k+1} f(x_k)}{1-\Delta_N },

@@

\frac{f'(x_k)}{f(x_k)} = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.

@@

Since $f(x)$ is log-concave, symmetric, and differentiable, we may write $f(x) = e^{c(x)}$ where $c(x)$ is concave, symmetric, and differentiable. We have 
\frac{f'(x)}{f(x)} = c'(x), \quad x\in \mathbb R,
which is anti-symmetric, non-negative for $x<0$,  non-positive for $x>0$, and non-increasing since $c(x)$ is concave. Therefore, if $c'(x_i) = c'(x_{i+1})$ for some $i =1,\ldots,N-1$, then either (1) $x_i = x_{i+1}$ or (2) $c'(x)$ is the zero function. Since (2) violates the assumption that $f(x)$ is a density function, we conclude that $c'(x)$ is an injection. 
As a result, \eqref{eq:gradient_zero} is satisfied if and only if $x_1 = \ldots = x_N$. For odd $N$ and $x_1=\ldots =x_N$, the LHS of \eqref{eq:lem_bound_intervals_proof} equals $\eta(x_1) = \max_i \eta(x_i)$ hence the statement holds. For even $N$ and any constant $d$, the limit of the LHS of \eqref{eq:lem_bound_intervals_proof} as $(x_1,\ldots,x_N)\rightarrow (d,\ldots,d)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:lem_bound_intervals_proof} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:lem_bound_intervals_proof} is maximized at the borders. That is, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches $\pm \infty$,  or $\pm \epsilon$. As we assumed $x_1 \geq \ldots \geq x_N$, if $x_i = x_{i+1}$ for some $i$ than their contribution to \eqref{eq:lem_bound_intervals_proof} is zero and thus this case reduces to the case $n= N-2$. A similar reduction holds if $x_N, x_{N-1} \to -\infty$, $x_1, x_2 \to \infty$, or  $x_i, x_{i+1}$ for some $i$. It is left to consider the cases:

@@

\lim_{x_N \to -\infty} \frac{  \delta_N^2} 
{\Delta_N \left(1- \Delta_N \right) }
 = \frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}f(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k) \right) \left(1- \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k)  \right)} ,

@@

which is smaller than $\max_i \eta_i(x_i)$ by the induction hypothesis. Under case (2) we have

@@

&  \lim_{x_1 \to \infty}
\frac{ \delta_N} 
{\Delta_N\left(1- \Delta_N \right) } \\

@@

& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}f(x_k) \right)} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} F(x_k) \right) \left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} F(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}f(x'_m) \right)^2 } 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m})  \right) },

@@

where $x'_{m} = x_{m+1}$. The last expression is also smaller than $\max_i \eta_i(x_i)$ by the induction hypothesis.  \QEDA \\

\subsection{Proof of Theorem~\ref{thm:adpative_lower_bound}
\label{proof:thm:adpative_lower_bound}
}

We first prove the following lemma:
\begin{lem} \label{lem:fisher_bound}
Let $X$ be a random variable with a symmetric, log-concave, and continuously differentiable density function $f(x)$ such that $\eta(x)$ is unimodal. For a Borel measurable $A$ set, 
\[
M(X) = \begin{cases} 1,& X \in A, \\
-1, & X \notin A.
\end{cases}
\]
Then the Fisher information of $M$ with respect to $\theta$ is bounded from above by $\eta(0)$.
\end{lem}

\subsubsection*{Proof of Lemma~\ref{lem:fisher_bound}}
When $f(x)$ is the normal density function, this lemma follows from \cite[Thm. 3]{Barnes2018}. The proof below is based on a different techique than in \cite{Barnes2018}, and is valid for any log-concave symmetric density satisfying Assumption~\ref{assump:failure_rate}. \\

The Fisher information of $M$ with respect to $\theta$ is given by
\begin{align}
I_\theta & =  \mathbb E \left[ \left( \frac{d}{d\theta} \log P\left( M | \theta \right) \right)^2 |\theta \right] \nonumber \\
& = \frac{ \left(\frac{d}{d\theta} P(M=1|\theta) \right)^2}{P(M=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(M=-1|\theta) \right)^2} {P(M=-1| \theta)} \nonumber \\
& =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(M=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(M=-1| \theta) } \nonumber \\ 
& \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(M=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(M=-1| \theta) } \nonumber \\ 
& = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(M=1 | \theta) \left(1-P(M=1|\theta) \right)  }, \nonumber \\
& = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
\end{align}
where differentiation under the integral sign in $(a)$ is possible since $f(x)$ is differentiable with continuous derivative $f'(x)$. Regularity of the Lebesgue measure implies that for any $\epsilon>0$, there exists a finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that 
\[
\int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon,
\]
which implies that for any $\epsilon' > 0$, the set $A$ in \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of disjoint intervals without increasing $I_\theta$ by more than $\epsilon'$. It is therefore enough to proceed in the proof assuming that $A$ is of the form
\[
A = \cup_{j=1}^k (a_j,b_j),
\]
with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
\begin{align*}
\mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
& = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
\end{align*}
so \eqref{eq:lem_fisher_bound_proof1} can be rewritten as
\begin{align}
& =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
{ \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
& \times \frac {1} 
{1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
\label{eq:lemma_J}
\end{align}
It follows from Lemma~\ref{lem:bound_intervals_delta} that for any $\theta \in \mathbb R$ and any choice of the intervals endpoints, \eqref{eq:lemma_J} is smaller than $4f^2(0)$.  \QEDA \\

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$, we first note that without loss of generality, each message $B_i$ can is of the form
\begin{equation}
\label{eq:general_messages}
B_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \mathbb R$ is a Borel measurable set. 
%Indeed, any measurable function $M(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
Consider the conditional distribution $P({B^n|\theta})$ of $B^n$ given $\theta$. We have 
\begin{align}
P\left( B^n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B^{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(B_i =1 | \theta, B^{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$, so that the Fisher information of $B^n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(B^n) = \sum_{i=1}^n I_\theta (B_i|B^{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B^{i-1})$ is the Fisher information of the distribution of $B_i$ given $B^{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (B_i|B^{i-1}) \leq 4f^2(0)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\mathbb E \left( \theta_n - \theta \right)^2 &  \geq \frac{1}{ \mathbb E I_\theta(B^n) + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}.
\end{align*}

@@

\subsubsection*{Proof of (i)}
%Specifically, (i) in Theorem~\ref{thm:sgd} 
Consider the following simplified version of \cite[Thm. 4]{polyak1992acceleration}:
\begin{thm}{\cite[Thm. 4]{polyak1992acceleration}} \label{thm:polyak_juditsky}
Let 
\[
X_i = \theta + Z_i,\quad i=1,\ldots,n,
\]
where the $Z_i$s are i.i.d. with zero means and finite variances. Define
\begin{align}
\begin{split}
\theta_i & = \theta_{i-1} + \gamma_i \varphi(X_i - \theta_{i-1}), \\
\bar{\theta}_n & = \frac{1}{n} \sum_{i=0}^{n-1} \theta_i, 
\end{split}
\label{eq:Polyak_Juditsky_alg}
\end{align}
where in addition, assume the following: 
\begin{enumerate}
\item[(i)] There exits $K_1$ such that $\left| \varphi(x) \right| \leq K_1(1+|x|)$ for all $x\in \mathbb R$.
\item[(ii)] The sequence $\left\{ \gamma_i \right\}_{i=1}^\infty$ satisfies conditions \eqref{eq:conditions1}.
\item[(iii)] The function $\psi(x) \triangleq \ex{ \varphi(x+Z_1)}$ is differentiable at zero with $\psi'(0)>0$, and satisfies $\psi(0)=0$ and $x\psi(x) >0$ for all $x\neq 0$.
Moreover, assume that there exists $K_2$ and $0<\lambda \leq 1$ such that
\begin{equation}
\label{eq:Polyak_Juditsky_cond3}
\left| \psi(x) - \psi'(0)x \right|\leq K_2 |x|^{1+\lambda}.
\end{equation}
\item[(iv)] The function 
$\chi(x) \triangleq \ex{\varphi^2(x+Z_1)}$ is continuous at zero. 
\end{enumerate}
Then $\bar{\theta}_n \rightarrow \theta$ almost surely and $ \sqrt{n}({\theta}_n - \theta)$ converges in distribution to $\Ncal(0,V)$, where
\[
V = \frac{ \chi(0)} {\psi'^2(0)}. 
\]
\end{thm}

Using the notation above, we set $\varphi(x) = \sgn(x)$ and $Z_i = X_i - \theta$. We have that $\chi(x) = \mathbb E \sgn^2(x+Z_1) = 1$, so $\chi(0) = 1$. In addition,
\begin{align*}
\psi(x) & = \ex{ \sgn(x+ Z_1) }= \int_{-\infty}^\infty \sgn(x+z) f(z) dz \\
& = \int_{-x}^\infty f(z) dz -\int_{-\infty}^{-x} f(z) dz. 
\end{align*}
Using the symmetry of $f(x)$ around zero, it follows that $\psi'(x) = 2f(x)$ and thus $\psi'(0) = 2f(0)$. It is now easy to verify that the rest of the conditions in Theorem~\ref{thm:polyak_juditsky} are fulfilled for any $\lambda > 0$. Since 
\[
\frac{\chi(0)}{\psi'^2(0)} = \frac{1}{4 f^2(0)} = \frac{1}{\eta(0)},
\]
it follows from Theorem~\ref{thm:polyak_juditsky} that
\[
\sqrt{n}\left({\theta}_n-\theta \right) \overset{d}{\to} \Ncal \left(0, 1/ \eta(0)\right). 
\]

>>>>

\subsection{Proof of Theorem~\ref{thm:adpative_lower_bound}
\label{proof:thm:adpative_lower_bound}
}
Consider the following two Lemmas:
\begin{lem} \label{lem:bound_intervals}
Let $f(x)$ be log-concave, symmetric, and differentiable density function such that Assumption~1 holds. For any $x_1 \geq \ldots \geq x_n \in \R$ and $\delta>0$,
\begin{equation}
\frac{ \left| \sum_{k=1}^n (-1)^{k+1} f(x_k) \right|^2 }
{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) \left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
\leq  4f^2(0). 
\label{eq:lem_bound_intervals}
\end{equation}
\end{lem}
\begin{lem} \label{lem:fisher_bound}
Let $X$ be a random variable with a symmetric, log-concave, and continuously differentiable density function $f(x)$ such that Assumption~1 holds. For a Borel measurable $A$ set, 
\[
M(X) = \begin{cases} 1,& X \in A, \\
-1, & X \notin A.
\end{cases}
\]
Then the Fisher information of $M$ with respect to $\theta$ is bounded from above by $\eta(0)$.
\end{lem}

Lemma~\ref{lem:bound_intervals} is obtained as the special case $\delta = 0$ of lemma \ref{lem:bound_intervals_delta}. 
We now prove Lemma~\ref{lem:fisher_bound}.
\subsubsection*{Proof of Lemma~\ref{lem:fisher_bound}}
When $f(x)$ is the normal density function, this lemma follows from \cite[Thm. 3]{Barnes2018}. The proof below is based on a different techique than in \cite{Barnes2018}, and is valid for any log-concave symmetric density satisfying Assumption~\ref{assump:failure_rate}. \\

The Fisher information of $M$ with respect to $\theta$ is given by
\begin{align}
I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( M | \theta \right) \right)^2 |\theta } \nonumber \\
& = \frac{ \left(\frac{d}{d\theta} P(M=1|\theta) \right)^2}{P(M=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(M=-1|\theta) \right)^2} {P(M=-1| \theta)} \nonumber \\
& =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(M=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(M=-1| \theta) } \nonumber \\ 
& \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(M=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(M=-1| \theta) } \nonumber \\ 
& = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(M=1 | \theta) \left(1-P(M=1|\theta) \right)  }, \nonumber \\
& = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
\end{align}
where differentiation under the integral sign in $(a)$ is possible since $f(x)$ is differentiable with continuous derivative $f'(x)$. Regularity of the Lebesgue measure implies that for any $\epsilon>0$, there exists a finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that 
\[
\int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon,
\]
which implies that for any $\epsilon' > 0$, the set $A$ in \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of disjoint intervals without increasing $I_\theta$ by more than $\epsilon'$. It is therefore enough to proceed in the proof assuming that $A$ is of the form
\[
A = \cup_{j=1}^k (a_j,b_j),
\]
with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
\begin{align*}
\mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
& = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
\end{align*}
so \eqref{eq:lem_fisher_bound_proof1} can be rewritten as
\begin{align}
& =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
{ \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
& \times \frac {1} 
{1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
\label{eq:lemma_J}
\end{align}
It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in \R$ and any choice of the intervals endpoints, \eqref{eq:lemma_J} is smaller than 
\[
\max_{t \in \{a_j,b_j, j=1,\ldots,k\} } 4f^2(t) \leq 4 f^2(0), 
\]
where the last transition is due to Assumption~1. 
\QEDA \\

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$, we first note that without loss of generality, each message $B_i$ can is of the form
\begin{equation}
\label{eq:general_messages}
B_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \R$ is a Borel measurable set. 
%Indeed, any measurable function $M(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
Consider the conditional distribution $P({B^n|\theta})$ of $B^n$ given $\theta$. We have 
\begin{align}
P\left( B^n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B^{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(B_i =1 | \theta, B^{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$, so that the Fisher information of $B^n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(B^n) = \sum_{i=1}^n I_\theta (B_i|B^{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B^{i-1})$ is the Fisher information of the distribution of $B_i$ given $B^{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (B_i|B^{i-1}) \leq 4f^2(0)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B^n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}.
\end{align*}
\QEDA

\subsection{Isoperimetric Lemma
\label{sec:bound_intervals_delta} }
The following lemma is used in the proof Theorems~\ref{thm:adpative_lower_bound}, \ref{thm:LAN1}, and \ref{thm:non_existence}.

\begin{lem} \label{lem:bound_intervals_delta}
Let $f(x)$ be log-concave, symmetric, and differentiable density function such that Assumption~1 holds. For any $x_1 \geq \ldots \geq x_n \in \R$ and $\delta>0$,
\begin{equation}
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left( \sum_{i=1}^n (-1)^{i+1} F(x_i) \right) \left(1- \sum_{k=1}^n (-1)^{i+1} F(x_i) \right) } 
\leq  \max_{i} \eta_\delta(x_i),
\label{eq:lem_bound_intervals_delta}
\end{equation}
where 
\[
\eta_\delta(x) \triangleq \frac{  \left( f(x) \right)^{2+\delta}}{\left(F(x)(1-F(x))\right)^{1+\delta}}. 
\]
\end{lem}

\subsubsection*{
Proof of Lemma~\ref{lem:bound_intervals_delta}}

@@

\Delta_n \triangleq \Delta_n(x_1,\ldots,x_n) \triangleq  \sum_{k=1}^n (-1)^{k+1} F(x_k). 

@@

Note that 

@@

\eta_\delta(x) =  \frac{  \left(\delta_1(x) \right)^{2+\delta}}{ (\Delta_1(x))^{1+\delta}
(1-\Delta_1(x))^{1+\delta}}. 

@@

Since $f(x)$ is log-concave, symmetric, and differentiable, we may write $f(x) = e^{c(x)}$ where $c(x)$ is concave, symmetric, and differentiable, with a non-increasing derivative $c'(x)$. We first prove the lemma under the assumption that 
c'(x) \triangleq \frac{f'(x)}{f(x)}, 
is an injection, or, equivalently, that $c(x)$ is strictly decreasing for all $x\in \R$. \par
We use induction on $n \in \N$ to show that the LHS of \eqref{eq:lem_bound_intervals_delta} is bounded from above by $\max_i \eta_\delta(x_i)$. The case $n=1$ is trivial. Assume that 

@@

{\left(\Delta_n\left(1- \Delta_n \right) \right)^{1+\delta} } \leq \max_i \eta_{\delta}(x_i)

@@

The maximal value of the LHS of \eqref{eq:lem_bound_intervals_proof} is attained for the same $(x_1,\ldots,x_N) \in \R^N$ that attains the maximal value of 

@@

@@

The partial derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by

@@

\frac{\partial  g}{\partial x_k} = \frac{(2+\delta) (-1)^{k+1} f'(x_k)}{\delta_N} -\frac{(1+\delta) (-1)^{k+1} f(x_k)}{\Delta_N } + \frac{(1+\delta)(-1)^{k+1} f(x_k)}{1-\Delta_N },

@@

c'(x_k) = \frac{f'(x_k)}{f(x_k)} = \frac{1+\delta}{2+\delta} \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.

@@

Since we assumed that $c'(x)$ is an injection, \eqref{eq:gradient_zero} is satisfied if and only if $x_1 = \ldots = x_N$. For odd $N$ and $x_1=\ldots =x_N$, the LHS of \eqref{eq:lem_bound_intervals_proof} equals $\eta(x_1) = \max_i \eta(x_i)$ hence the statement holds. For even $N$ and any constant $d$, the limit of the LHS of \eqref{eq:lem_bound_intervals_proof} as $(x_1,\ldots,x_N)\rightarrow (d,\ldots,d)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:lem_bound_intervals_proof} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:lem_bound_intervals_proof} is maximized at the borders. That is, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches $\pm \infty$,  or $\pm \epsilon$. As we assumed $x_1 \geq \ldots \geq x_N$, if $x_i = x_{i+1}$ for some $i$ than their contribution to \eqref{eq:lem_bound_intervals_proof} is zero and thus this case reduces to the case $n= N-2$. A similar reduction holds if $x_N, x_{N-1} \to -\infty$, $x_1, x_2 \to \infty$, or  $x_i, x_{i+1}$ for some $i$. It is left to consider the cases:

@@

@@

@@

@@

@@

where $x'_{m} = x_{m+1}$ for $m=1,\ldots, N-1$. The last expression is also smaller than $\max_i \eta_\delta (x_i)$ by the induction hypothesis.  \par
It is left to consider the case where $c'(x)$ is not necessarily strictly decreasing. We approximate $c(x)$ using another concave symmetric function whose derivative is always negative except at the origin. For $\alpha>0$ consider the function  $f_\alpha(x) = \kappa(\alpha) e^{\sgn(c(x))|c(x)|^{1+\alpha}}$, where $\kappa(\alpha)$ is chosen such that $f_\alpha(x)$ is a probability density function. Then $c_\alpha(x)$ is concave, symmetric, and differentiable with
c_\alpha'(x) \triangleq \frac{f'_\alpha(x)}{f_\alpha(x)} = (1+\alpha)|c(x)|^{\alpha} c'(x). 
Now $c_\alpha'(x)$ is non-increasing since it is the derivative of a concave function. It can be verified that it is non-constant on any interval (because $c(x)$ is non-constant and $c'(x)$ is non-increasing), and hence an injection. It follows from the first part of the proof that, for any $\alpha>0$,
\label{eq:proof:lem:bound_intervals}
\frac{(\delta_{n,\alpha})^2}{\Delta_{n,\alpha}(1-\Delta_{n,\alpha})} \leq \max_i \eta_{\alpha}(x_i),
where 
\delta_{n,\alpha} \triangleq  \sum_{k=1}^{n} (-1)^{k+1} f_{\alpha}(x_k),
\Delta_{n,\alpha} \triangleq \sum_{k=1}^n (-1)^{k+1} F_{\alpha}(x_k), 
\eta_{\delta,\alpha}(x) \triangleq \frac{(f_\alpha(x))^{2+\delta}}{\left(F_{\alpha}(x)(1-F(x)) \right)^{1+\delta}}. 
The proof is completed by noting that 
\lim_{\alpha \to 0} \frac{(\delta_{n,\alpha})^{2+\delta} }{ \left(\Delta_{n,\alpha}(1-\Delta_{n,\alpha})\right)^{1+\delta}}  = \frac{(\delta_{n})^{2+\delta }}{\left(\Delta_{n}(1-\Delta_{n}) \right)^{1+\delta}},  
and, since the maximum is over a finite set,
\lim_{\alpha \to 0}  \max_i \eta_{\delta,\alpha}(x_i)  = \max_i\eta_\delta(x_i).

%Our goal is to construct an approximation to $f(x)$ using a log-concave, symmetric, and differentiable function $\hat{f}(x)$ such that $\hat{f}'(x)/\hat{f}(x)$ is strictly decreasing. We achieve such an approximation by modifying $f(x)$ over intervals over which $c'(x)$ is a constant. Assume first $c'(x)$ is only a constant over an interval $(a_1,b_1)\subset \R$ containing $x_1>0$ and on the mirror image of this interval around $x=0$. Over this interval, $f(x)$ is necessarily of the form $f(x) = e^{-c_1|x| + d_1}$ for some $c_1\geq 0$ and $d_1$. % and For $\alpha>0$, consider
%\[
%F_1(x,\alpha) \triangleq  F(a_1) + \frac{1}{c_1} e^{-c_1 (|x|^{1+\alpha} +d_1}, \quad x \in (a_1,b_1),
%\]
%and 
%\[
%f_1(x,\alpha) \triangleq  F_1'(x,\alpha) =  -\sgn(x)(1+\alpha)|x|^\alpha e^{-c_1 |x|^{1+\alpha} + d_1}, \quad x \in (a_1,b_1).
%\]

@@

\subsubsection*{Proof of (i)}
%Specifically, (i) in Theorem~\ref{thm:sgd} 
Consider the following simplified version of \cite[Thm. 4]{polyak1992acceleration}:
\begin{thm}{\cite[Thm. 4]{polyak1992acceleration}} \label{thm:polyak_juditsky}
Let 
\[
X_i = \theta + Z_i,\quad i=1,\ldots,n,
\]
where the $Z_i$s are i.i.d. with zero means and finite variances. Define
\begin{align}
\begin{split}
\theta_i & = \theta_{i-1} + \gamma_i \varphi(X_i - \theta_{i-1}), \\
\bar{\theta}_n & = \frac{1}{n} \sum_{i=0}^{n-1} \theta_i, 
\end{split}
\label{eq:Polyak_Juditsky_alg}
\end{align}
where in addition, assume the following: 
\begin{enumerate}
\item[(i)] There exits $K_1$ such that $\left| \varphi(x) \right| \leq K_1(1+|x|)$ for all $x\in \R$.
\item[(ii)] The sequence $\left\{ \gamma_i \right\}_{i=1}^\infty$ satisfies conditions \eqref{eq:conditions1}.
\item[(iii)] The function $\psi(x) \triangleq \ex{ \varphi(x+Z_1)}$ is differentiable at zero with $\psi'(0)>0$, and satisfies $\psi(0)=0$ and $x\psi(x) >0$ for all $x\neq 0$.
Moreover, assume that there exists $K_2$ and $0<\lambda \leq 1$ such that
\begin{equation}
\label{eq:Polyak_Juditsky_cond3}
\left| \psi(x) - \psi'(0)x \right|\leq K_2 |x|^{1+\lambda}.
\end{equation}
\item[(iv)] The function 
$\chi(x) \triangleq \ex{\varphi^2(x+Z_1)}$ is continuous at zero. 
\end{enumerate}
Then $\bar{\theta}_n \rightarrow \theta$ almost surely and $ \sqrt{n}({\theta}_n - \theta)$ converges in distribution to $\Ncal(0,V)$, where
\[
V = \frac{ \chi(0)} {\psi'^2(0)}. 
\]
\end{thm}

Using the notation above, we set $\varphi(x) = \sgn(x)$ and $Z_i = X_i - \theta$. We have that $\chi(x) = \ex{ \sgn^2(x+Z_1) }= 1$, so $\chi(0) = 1$. In addition,
\begin{align*}
\psi(x) & = \ex{ \sgn(x+ Z_1) }= \int_{-\infty}^\infty \sgn(x+z) f(z) dz \\
& = \int_{-x}^\infty f(z) dz -\int_{-\infty}^{-x} f(z) dz. 
\end{align*}
Using the symmetry of $f(x)$ around zero, it follows that $\psi'(x) = 2f(x)$ and thus $\psi'(0) = 2f(0)$. It is now easy to verify that the rest of the conditions in Theorem~\ref{thm:polyak_juditsky} are fulfilled for any $\lambda > 0$. Since 
\[
\frac{\chi(0)}{\psi'^2(0)} = \frac{1}{4 f^2(0)} = \frac{1}{\eta(0)},
\]
it follows from Theorem~\ref{thm:polyak_juditsky} that
\[
\sqrt{n}\left({\theta}_n-\theta \right) \overset{d}{\to} \Ncal \left(0, 1/ \eta(0)\right). 
\]
